Lesson 3
========================================================

***

### What to Do First?
Notes:

***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
getwd()
setwd('~/Udacity/Data Analysis with R/lesson3/')
list.files()
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
names(pf)
```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
#install.packages('ggplot2')
library(ggplot2)

names(pf)
qplot(data = pf, x = dob_day) +
  scale_x_discrete(breaks=1:31)

```

***

#### What are some things that you notice about this histogram?
Response: The count of dob_day = 1 seems to be a large outlier.  The count of dob_day = 31 seems to be relatively small, but not ever month has 31 days, so this seems reasonable.

***

### Moira's Investigation
Notes: Users underestimated their audience, guessing a quarter of the actual audience size

***

### Estimating Your Audience Size
Notes: 

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response: A status about going down a YouTube video rabbit hole

#### How many of your friends do you think saw that post?
Response: 100

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response: 20

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:

```{r Faceting}
ggplot(data = pf, aes(x = dob_day)) +
  geom_histogram() +
  scale_x_discrete(breaks = 1:31) +
  facet_wrap(~dob_month)
```

#### Let’s take another look at our plot. What stands out to you here?
Response: It appears this high count of dob_day = 1 occurs when dob_month = 1.  This is in line with the assertion that the default birthday presented to the user is January 1st.

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes: 
#### Which case do you think applies to Moira’s outlier?
Response: Bad data about an extreme case

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
ggplot(data = pf, aes(x = friend_count)) +
  geom_histogram()
```

#### How is this plot similar to Moira's first plot?
Response: A large portion of the data is in a small bin

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}

```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender)

by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response: women

#### What's the difference between the median friend count for women and men?
Response: 22

#### Why would the median be a better measure than the mean?
Response: Less susceptible to outliers

***

### Tenure
Notes:

```{r Tenure}
ggplot(aes(x = tenure), data = pf) + 
   geom_histogram(binwidth = 30, color = 'black', fill = '#099DD9')
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
ggplot(aes(x = tenure/365), data = pf) + 
  geom_histogram(binwidth = 0.25, color = 'black', fill = '#F79420') +
  scale_x_continuous(breaks = seq(1, 7, 1), lim = c(0, 7)) +
  xlab('Number of years using Facebook') + 
  ylab('Number of users in sample')
  
```

***

### Labeling Plots
Notes:

```{r Labeling Plots}

```

***

### User Ages
Notes:

```{r User Ages}
ggplot(data = pf, aes(x = age)) +
  geom_histogram()
```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:

***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}
library(gridExtra) 
library(ggplot2)

p1 = ggplot(data = pf, aes(x = friend_count)) +
      geom_histogram()

p2 = p1 + scale_x_log10()

p3 = p1 + scale_x_sqrt()

grid.arrange(p1, p2, p3, ncol = 1)
```

***


### Frequency Polygons

```{r Frequency Polygons}
ggplot(aes(x = www_likes, y = ..count../sum(..count..)), data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender)) + 
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  scale_x_log10() + 
  xlab('www_likes') + 
  ylab('Percentage of users with that many www_likes')
```

***

### Likes on the Web
Notes:

```{r Likes on the Web}
by(pf$www_likes, pf$gender, sum)
```


***

### Box Plots
Notes:

```{r Box Plots}
ggplot(data = subset(pf, !is.na(gender)), aes(x=gender, y=friend_count)) +
  geom_boxplot()
```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.
Notes: This method removes data, changing the percentile calculations
```{r}
ggplot(data = subset(pf, !is.na(gender)), aes(x=gender, y=friend_count)) +
  geom_boxplot() +
  scale_y_continuous(limits=c(0,1000))
```

Notes: This method does not remove data, and simply zooms the graph
```{r}
ggplot(data = subset(pf, !is.na(gender)), aes(x=gender, y=friend_count)) +
  geom_boxplot() +
  coord_cartesian(ylim=c(0,1000))
```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
ggplot(data = subset(pf, !is.na(gender)), aes(x=gender, y=friend_count)) +
  geom_boxplot() +
  coord_cartesian(ylim=c(0,250))

by(pf$friend_count, pf$gender, summary)
```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}
ggplot(data = subset(pf, !is.na(gender)), aes(x=gender, y=friendships_initiated)) +
  geom_boxplot() +
  coord_cartesian(ylim=c(0,150))


by(pf$friendships_initiated, pf$gender, summary)
```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}
pf$mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)

sum(pf$mobile_check_in == 1)/length(pf$mobile_check_in)
```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!