Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')

qplot(x = age, y = friend_count, data = pf)
```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point() +
  xlim(13, 90)

summary(pf$age)
```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_jitter(alpha = 1/20) +
  xlim(13, 90)
```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}
?coord_trans
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) +
  xlim(13, 90) +
  coord_trans(y = "sqrt")
```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) +
  xlim(13, 90) +
  coord_trans(y = "sqrt")
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
install.packages('dplyr')
library(dplyr)

age_groups <- group_by(pf, age)
pf.fc_by_age <- summarise(age_groups,
                          friend_count_mean = mean(friend_count),
                          friend_count_median = median(friend_count),
                          n = n())
pf.fc_by_age <- arrange(pf.fc_by_age, age)

pf.fc_by_age_chained <- pf %>% group_by(age) %>% 
                        summarise(friend_count_mean = mean(friend_count),
                                  friend_count_median = median(friend_count),
                                  n = n()) %>%
                        arrange(age)
```

Create your plot!

```{r Conditional Means Plot}
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) + 
  geom_line()
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  coord_cartesian(xlim = c(13, 70), ylim = c(0, 1000)) +
  geom_point(alpha = 1/20, 
             position = position_jitter(h = 0),
             color = "orange") +
  geom_line(stat = "summary", fun.y = mean) +
  geom_line(stat = "summary", fun.y = quantile, probs = 0.1,
            linetype = 2,
            color = "blue") +
  geom_line(stat = "summary", fun.y = quantile, probs = 0.5,
            color = "blue") +
  geom_line(stat = "summary", fun.y = quantile, probs = 0.9,
            linetype = 2,
            color = "blue")
```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor.test(pf$age, pf$friend_count)
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(pf[pf$age <= 70,], cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) +
  geom_point(alpha = 0.05) +
  coord_cartesian(xlim = c(0, quantile(pf$www_likes_received, 0.95)), 
                  ylim = c(0, quantile(pf$likes_received, 0.95))) +
  geom_smooth(method = "lm", color = "red")
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) +
  geom_point(alpha = 0.05) +
  coord_cartesian(xlim = c(0, quantile(pf$www_likes_received, 0.95)), 
                  ylim = c(0, quantile(pf$likes_received, 0.95))) +
  geom_smooth(method = "lm", color = "red")
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
pf.subset = subset(pf, www_likes_received > quantile(www_likes_received, 0.95) & likes_received > quantile(likes_received, 0.95))
cor.test(pf.subset$www_likes_received, pf.subset$likes_received)
cor.test(pf$www_likes_received, pf$likes_received)
```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
data(Mitchell)
?Mitchell
```

Create your plot!

```{r Temp vs Month}
ggplot(aes(x = Month, y = Temp), data = Mitchell) +
  geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

0

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

0.05747063 

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp)
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(aes(x = Month, y = Temp), data = Mitchell) +
  geom_point() +
  scale_x_discrete(breaks = seq(0, 203, 12))
```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
pf$age_with_months <- pf$age + (12 - pf$dob_month)/12
```

***

### Age with Months Means

```{r Age with Months Means}
pf.fc_by_age_months <- pf %>% group_by(age_with_months) %>% 
                        summarise(friend_count_mean = mean(friend_count),
                                  friend_count_median = median(friend_count),
                                  n = n()) %>%
                        arrange(age_with_months)

head(pf.fc_by_age_months)
```

Programming Assignment
```{r Programming Assignment}
ggplot(aes(x = age_with_months, y = friend_count_mean), data = pf.fc_by_age_months[pf.fc_by_age_months$age_with_months < 71,]) +
  geom_line()
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

